---
title: "Using wordVectors to Uncover the 'Folk' in _Billboard Magazine_"
author: "Jessica Dauterive"
date: "4/23/2017"
output: html_document
---
```{r setup, include=FALSE}
library(devtools)
library(wordVectors)
library(magrittr)
library(tokenizers)
library(ggplot2)
library(dplyr)
library(knitr)
library(plotly)
billboard = read.vectors("billboards_vectors.bin")
```

This essay will problematize distinctions between genres in _Billboard Magazine_ by taking a large sample of issues through time (1930s-2010s), and using wordVectors to find similarities between particular words. Here, I plan to explore the ways that "folk" musics are talked about, by searching for the words that have come to reference traditional music traditions--"country" and "blues." Surely, country and blues music both stem from southern traditions, a soundscape that is often mis-remembered as isolated from popular culture and segregated, despite shared communities more united by being laborers than divided by race and connected with the rest of nation through technologies of transmission and transportation. Yet, once record companies became interested in recording "southern" music, musicians were separated from one another largely for marketing purposes--black artists were packaged as "race" records and white artists as "hillbilly" records and were then easy to sell to retail stores in a segregated America. This text analysis reflects this complicated history around recording American folk music that not only defies the color line, but also suggests a larger sphere of musical influence than what is typically understood as "traditional" roots music.[^1]

```{r, echo=FALSE, warning=FALSE}
genres = billboard[[c("blues","country"),average=F]]

#take 4500 most common words
countryness_and_bluesiness = billboard[1000:4500,] %>% 
  cosineSimilarity(genres)

#filter to top 150
c_b_filter = countryness_and_bluesiness[
  rank(-countryness_and_bluesiness[,1])<150 |
  rank(-countryness_and_bluesiness[,2])<150,
  ] %>% 
  as.data.frame() %>% 
  tibble::rownames_to_column()

#plot
filter_graph <- c_b_filter %>%
  ggplot(aes(x = blues, y = country)) +
  geom_text(aes(label = rowname)) 
ggplotly(filter_graph)
```

Plotting the top 150 words with similarities between "country" and "blues" reveals at first glance that there are very few words that have a close relationship to both. Rather, they are clusetered as words that are either more similar to country or more similar to blues, perhaps suggesting that there is actually less overlap between these genres than suggested above. In fact, zooming in on the country chunk (it is an interactive graph, so draw a square with your cursor around the segment you wish you view) reveals words like "depression", "jobbers", "wheeling", "flood", and "hills", playing into ideas about country musicians as iternerant, working class folks, and especially reflecting the way that folklorists "discovered" these musicians during the Great Depression. However, zooming in on the "blues" chunk reveals some suprising terms.

First, you find words like "latin", "spanish", and "rumba". You also find "harlem", which certainly does not fit the understanding of blues music as a southern tradition, and suggests that _Billboard_ might have talked about blues in close alignment with jazz. (In fact, running similarities between "blues" and "jazz" does reveal a very close positive relationship, and has its own implications for how race and geography (de)construct genre that are beyond the scope of this essay.) However, there are other words in this cluster that raise more quesions about the relationship between country and blues. "Negro" is displayed with about a .3 similarity to "blues," but also a .1 similarity to "country." Moreover, "cowboy" and "hillbilly" both skew more towards "blues" than "country." This suggests that the early origins of blues music (and perhaps later understandings, too) reflect shared musical traditions between black and white musicians, and brings attention to the obfuscation of black cowboys from historical memory.[^2] Another interesting word appears, "Hawaiian," which seems surprising because Hawaiin music has such a strong connection to white, middle class leisure in the 1950s and 1960s. However, as historian John Troutman has recently uncovered, Hawaiian steel guitar had a much larger emphasis on blues music than was previously thought.[^3] What this graph begins to reveal, then, is a reconsideration of how American folk music memory has been structured, and what people and traditions were erased through the solidification of genre.

```{r, echo=FALSE,warning=FALSE}
genres3 = billboard[[c("blues","country","minstrel"),average=F]]

#take 4500 most common words
BMC_frame = billboard[2000:4500,] %>% 
  cosineSimilarity(genres3)

#filter to top 75
BMC_filter = BMC_frame[
  rank(-BMC_frame[,1])<150 |
  rank(-BMC_frame[,2])<150 |
  rank(-BMC_frame[,3])<150,
  ] %>% 
  as.data.frame() %>% 
  tibble::rownames_to_column()

minstrelsy_graph <- BMC_filter %>% 
  ggplot(aes(x = blues, y = country)) +
  geom_text(aes(label = rowname)) 
ggplotly(minstrelsy_graph)
```

Adding a third element to the analysis--"minstrel"--reveals some surprising findings that even further complicate our notion of folk music. First, you can see that a new cluster of words appears that are common to both "country" and "blues," and these words are those that relate to minstrelsy in a variety of ways. This raises a complicated question about the relationship between these genres (again, as genres, not musical traditions) and minstrelsy as perhaps their most common factor. This is one the questions at the root of the "Hearing the Americas" project, and seems to be supported by this graph which demonstrates, through shared words like "rube" and "comedian", that perhaps the popularity of comedic recordings of minstrelsy, vaudeville, blackface and other "novelties" might be the strongest connection betweeen the "roots" of these two genres. 

One final note--the word "folk" doesn't appear until analyzing the "country," "blues," and "minstrel" vectors together, and has a close positive relationship to both "blues" and "country," despite its contemporary connotations as a white, working class music. This can be read as suggestive of the disparate goals of folklorists and commerical recording companies to define and package an "American" musical tradition--folklorists sought traditional "folk" while record companies wanted "hits" that still relied on popular live traditions of big band, minstrelsy, and jazz. Other evidence to this point is the close relationship between early labels like "victor," "decca," and "okeh" to "blues," and their negative relationships to "country." There are surely more themes revealed in this analysis, and many more vectors that can be run to investigate the questions they raise. However, these two graphs can be used as a starting point to unpack the murkiness of genre distinction, and suggests a reconsideration of what--and who--we write about when we write about folk music.

[^1]: Historian Karl Hagstrom Miller articulates these ideas in depth in, Miller, _Segregating Sound: Inventing Folk and Pop Music in the Age of Jim Crow_ (Durhham: Duke University Press, 2010).
[^2]: http://www.npr.org/2010/12/05/131761541/we-ve-all-heard-cowboy-songs-but-who-were-the-cowboys
[^3]: http://www.oxfordamerican.org/magazine/item/1166-the-winsome-moan 
