---
title: "textAnalysisEssay"
author: "Jessica Dauterive"
date: "4/21/2017"
output: html_document
---
```{r setup, include=FALSE}
library(devtools)
library(wordVectors)
library(magrittr)
library(tokenizers)
library(ggplot2)
library(dplyr)
library(knitr)
library(plotly)
billboard = read.vectors("billboards_vectors.bin")
```

This essay will investigate how genre was spoken about in Billboard Magazine by taking a large sample of pages and years (1930s-2010s) and using wordVectors to find similarities between certain words. In this instance, I am curious how the blues was talked about in relation to country and in relation to jazz. I chose these comparisons because they are genres that became racially coded around the time that this corpus starts. I chose to use the entire run of Billboard (rather than choosing a selection) in order to capture the widest range of ways that these two genres were spoken about. What is revealed here, is perhaps not all surprising: jazz and blues are more closely related than blues and country. However, within these larger statements, there are interesting findings that suggest possibilities for further research on the process of genre in the music industry.

My hypothesis is that Billboard discussed genre more in terms of race than in terms of musical qualities. To do so, I will first exmamine the relationship between country and blues music, which shares similiar instrumentation and geographic associations. I will then compare blues and jazz music, that have different musical traditions and are typically divided between rural and urban areas. My hypothesis is that jazz and blues will have more in common than blues and country because of the way that genre become a racially coded tool of the music indsutry.



First, here is a graph of the words that share a similiarity between country and blues. As you can see from this chart, country and blues have a negative relationship, and this sample of the vector below indicates this relationship. 
```{r, echo=FALSE, warning=FALSE}
genres = billboard[[c("blues","country"),average=F]]

#take 4500 most common words
countryness_and_bluesiness = billboard[1000:4500,] %>% 
  cosineSimilarity(genres)

#filter to top 75
c_b_filter = countryness_and_bluesiness[
  rank(-countryness_and_bluesiness[,1])<150 |
  rank(-countryness_and_bluesiness[,2])<150,
  ] %>% 
  as.data.frame() %>% 
  tibble::rownames_to_column()

#plot
filter_graph <- c_b_filter %>%
  ggplot(aes(x = blues, y = country)) +
  geom_text(aes(label = rowname)) 
ggplotly(filter_graph)
```


This chart of the top 75 words (chosen for clarity and relevance) reveals several expected terms in common: waltz, tune, and bluebird all refer to country and blues, but skew towards blues. However, there are a few surprises. Although the term "negro" which would be expected to be more closely related to the blues does fall there, the term "hillbilly" has a negative relationship to country and a positive one to blues. There are other interesting pieces of the chart. There are various ethnicities, irish and latino, that might not always be considered with country and the blues, yet have a positive relationship to both. Also, the word "rumba" has a positive relationship. This seems suprising, since blues is usually considered to be a southern, black music.



Harlem has almost an equally positive relationship, which suggests that these two genres were spoken about together. With such close connections some of the more suprising terms (rumba, latin, gypsy) make a lot more sense, since they are often used to describe jazz. This indicates a problematic relationship between genre as a racial distinction as it formed in the 1930s and as a musical distinction, since blues music should have more in common with country (geographically and tranditionally) than jazz. What this begins to suggest is the inadequacy of defining music within genre distinctions.

These charts suggest further lines of inquiry. It would be interesting to compare instrumentation between genres to see how musical distinction plays into this. It would also be interesting to see how gender and value judgments might interfere with genre. It might also be useful to do a topic model on the genres themselves 

```{r, echo=FALSE,warning=FALSE}
genres3 = billboard[[c("blues","country","minstrel"),average=F]]

#take 4500 most common words
BMC_frame = billboard[1000:4500,] %>% 
  cosineSimilarity(genres3)

#filter to top 75
BMC_filter = BMC_frame[
  rank(-BMC_frame[,1])<50 |
  rank(-BMC_frame[,2])<50 |
  rank(-BMC_frame[,3])<50,
  ] %>% 
  as.data.frame() %>% 
  tibble::rownames_to_column()

minstrelsy_graph <- BMC_filter %>% 
  ggplot(aes(x = blues, y = country)) +
  geom_text(aes(label = rowname)) 
ggplotly(minstrelsy_graph)
```
